---
title: "Initial Analysis - grand center NDVI"
output: html_notebook
---

```{r setup, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(lme4)
library(multilevel)
library(sjPlot)
library(sjmisc)
library(broom)

df <- readRDS("../data/full_mlm_datset.rds")
```
Let's look at the variables of interest. These are crime index-total, crime index-violent, crime index-property, and green measures.
```{r}
skimr::skim(df$CRMCYTOTC)
skimr::skim(df$CRMCYPERC)
skimr::skim(df$CRMCYPROC)
skimr::skim(df$gndctr_ndvi_cbg_adj)
```
First, is there variation in the variables that is attributable to the differences between cities, the level 2 'groups'.
```{r}
t <- df[, c("CRMCYTOTC", "CRMCYPERC", "CRMCYPROC", "gndctr_ndvi_cbg_adj")]
mult.icc(as.data.frame(t), grpid= df$city_st)
```

Start with the intercept only models
```{r}
mod_1a <- lmer(CRMCYTOTC ~ 1 + (1|city_st), data = df)
mod_1b <- lmer(CRMCYPERC ~ 1 + (1|city_st), data = df)
mod_1c <- lmer(CRMCYPROC ~ 1 + (1|city_st), data = df)

summary(mod_1a)
summary(mod_1b)
summary(mod_1c)
```
So maybe at this point let's just look at the ndvi relationship. First as just random intercepts.
```{r}
mod_2a <- lmer(CRMCYTOTC ~ gndctr_ndvi_cbg_adj + (1|city_st), data = df)
mod_2b <- lmer(CRMCYPERC ~ gndctr_ndvi_cbg_adj + (1|city_st), data = df)
mod_2c <- lmer(CRMCYPROC ~ gndctr_ndvi_cbg_adj + (1|city_st), data = df)

summary(mod_2a)
summary(mod_2b)
summary(mod_2c)
```
Do these fit better than the intercept only models?
```{r}
anova(mod_1a, mod_2a)
anova(mod_1b, mod_2b)
anova(mod_1c, mod_2c)
```
Next, how about adding random slopes for NDVI. Now the NDVI slope varies by city, so some cities will have a stronger or weaker relationship between NDVI and crime.
```{r}
mod_3a <- lmer(CRMCYTOTC ~ gndctr_ndvi_cbg_adj + (gndctr_ndvi_cbg_adj|city_st), data = df)
mod_3b <- lmer(CRMCYPERC ~ gndctr_ndvi_cbg_adj + (gndctr_ndvi_cbg_adj|city_st), data = df)
mod_3c <- lmer(CRMCYPROC ~ gndctr_ndvi_cbg_adj + (gndctr_ndvi_cbg_adj|city_st), data = df)

summary(mod_3a)
summary(mod_3b)
summary(mod_3c)
```
Do the models with varying slopes provide a better fit?
```{r}
anova(mod_2a, mod_3a)
anova(mod_2b, mod_3b)
anova(mod_2c, mod_3c)
```
With allowing varying intercepts and slopes for NDVI we get an estimate for each city, let's plot those from the 3rd models.
```{r}
estmod3b <- coef(mod_3b)$city_st
ggplot(df, aes(gndctr_ndvi_cbg_adj, CRMCYPERC)) + geom_point(alpha = 0) + geom_abline(data = estmod3b, aes(slope = gndctr_ndvi_cbg_adj, intercept = `(Intercept)`), alpha = .2) + xlim(-10, 10) + ggtitle("Violent Crime") + geom_vline(aes(xintercept = mean(gndctr_ndvi_cbg_adj)), data = df)

estmod3c <- coef(mod_3c)$city_st
ggplot(df, aes(gndctr_ndvi_cbg_adj, CRMCYPROC)) + geom_point(alpha = 0) + geom_abline(data = estmod3c, aes(slope = gndctr_ndvi_cbg_adj, intercept = `(Intercept)`), alpha = .2) + xlim(-10, 10) + ggtitle("Property Crime") + geom_vline(aes(xintercept = mean(gndctr_ndvi_cbg_adj)), data = df)
```
With the NDVI measure grand mean centered we have 0 as the average across all block groups.(min of NDVI in cbg is `r min(df$gndctr_ndvi_cbg_adj)` and max of NDVI is `r  max(df$gndctr_ndvi_cbg_adj)`)
```{r, out.height= "220%"}
estmod3b %>% mutate(city_st = row.names(.)) %>% ggplot(aes(`(Intercept)`, gndctr_ndvi_cbg_adj)) + geom_point() + labs(x = "Intercept", y = "NDVI slope", title = "Violent Crime") 

estmod3c %>% mutate(city_st = row.names(.)) %>% ggplot(aes(`(Intercept)`, gndctr_ndvi_cbg_adj)) + geom_point() + labs(x = "Intercept", y = "NDVI slope", title = "Property Crime") 
```

