---
title: "Modeling - article 1"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
---
```{r setup, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(lme4)
library(multilevel)
library(sjPlot)
library(sjmisc)
library(broom)
library(merTools)
library(stargazer)
library(texreg)


city <- readRDS("../data/city_all.rds")
df <- readRDS("../data/full_mlm_datset.rds")
newclust <- city[, c(1, 68)]
df <- df %>% left_join(newclust)
df_na <- df %>% filter(!is.na(gndctr_MedHHinc000)& !is.na(gndctr_disad))
df_urban <- df %>% filter(Urban == 1)
df_urban_nona <- df_na %>% filter(Urban == 1)


```
I am going to run the models and then try to print that to a page, in landscape orientation.

First, with the raw, untransformed data. Using the data with NA removed as things like MedHHinc have some NAs.

```{r models 1, results='asis'}
mod_1 <- lmer(CRMCYPERC ~ 1 + (1|city_st), data = df_na)
mod_2 <- lmer(CRMCYPERC ~ ndvi_mean_cbg_adj + (1|city_st), data = df_na)
mod_3 <- lmer(CRMCYPERC ~ ndvi_mean_cbg_adj + (ndvi_mean_cbg_adj|city_st), data = df_na)
mod_4 <- lmer(CRMCYPERC ~ ndvi_mean_cbg_adj + MedHHinc_000 + disad + divindex + PcU18_whole + log_popden_cbg + (ndvi_mean_cbg_adj|city_st), data = df_na)
mod_5 <- lmer(CRMCYPERC ~ ndvi_mean_cbg_adj + MedHHinc_000 + disad + divindex + PcU18_whole + log_popden_cbg + PerCapitaOfficers1000 + PCgdp15 + factor(clust90_4) + rate_violentcrime + (ndvi_mean_cbg_adj|city_st), data = df_na)
```

\newpage
\blandscape
Landscape
```{r print1}
screenreg(list(mod_1, mod_2, mod_3, mod_4, mod_5))
```
\elandscape

\newpage
Now with the grand centered variables

```{r models 2}
mod_6 <- lmer(CRMCYPERC ~ 1 + (1|city_st), data = df_na)
mod_7 <- lmer(CRMCYPERC ~ gndctr_ndvi_cbg_adj + (1|city_st), data = df_na)
mod_8 <- lmer(CRMCYPERC ~ gndctr_ndvi_cbg_adj + (gndctr_ndvi_cbg_adj|city_st), data = df_na)
mod_9 <- lmer(CRMCYPERC ~ gndctr_ndvi_cbg_adj + gndctr_MedHHinc000 + gndctr_disad + gndctr_diver + gndctr_U18 + gndctr_logpopden + (gndctr_ndvi_cbg_adj|city_st), data = df_na)
mod_10 <- lmer(CRMCYPERC ~ gndctr_ndvi_cbg_adj + gndctr_MedHHinc000 + gndctr_disad + gndctr_diver + gndctr_U18 + gndctr_logpopden + gndctr_Police + PCgdp15 + factor(clust90_4) + gndctr_rt_violent + (gndctr_ndvi_cbg_adj|city_st), data = df_na)
```

\newpage
\blandscape

```{r print2, echo=FALSE}
screenreg(list(mod_6, mod_7, mod_8, mod_9, mod_10))
```
\elandscape

\newpage

## Interaction

Using grand mean centered variables, level 1 and level 2, let's add the interaction between Green and Climate.

```{r interactions, echo=FALSE}

mod_11 <- lmer(CRMCYPERC ~ gndctr_ndvi_cbg_adj*factor(clust90_4) + gndctr_MedHHinc000 + gndctr_disad + gndctr_diver + gndctr_U18 + gndctr_logpopden + gndctr_Police + PCgdp15 + gndctr_rt_violent + (gndctr_ndvi_cbg_adj|city_st), data = df_na)

screenreg(mod_11)
```

